---
title: "Mellow: a story"
author: "Steven van Wely"
date: "May the 7nd, 2024"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    self_contained: false
    source: https://github.com/Stevgit1/Computational-musicology-homework-portfolio/blob/main/index.Rmd
    theme:
      heading_font:
        google: 
          family: Rajdhani
          wght: 700
      base_font:
        google: Fira Sans
      code_font:
        google: Fira Mono
      bg: "#FFFFFF"
      fg: "#212529" 
      primary: "#2b2bee"
      secondary: "#39d7b8"
      success: "#39d7b8"
      danger: "#fa5577"
      warning: "#ffb14c"
      info: "#0cc7f1"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

```{r, setup}
library(tidyverse)
library(ggplot2)
library(plotly)
library(spotifyr)
library(compmus)
library(ggdendro)
library(heatmaply)
library(tidymodels)
library(cowplot)
library(glue)
library(gridExtra)
```

```{r, setup2}
# adding playlists
mellow_bars <- get_playlist_audio_features("", "37i9dQZF1DWT6MhXz0jw61")
hardcore_tracks <- get_playlist_audio_features("", "37i9dQZF1EIexrP9IBJXx5")
top100_tracks <- get_playlist_audio_features("", "37i9dQZEVXbNG2KDcFcKOF")
```

### Introduction

On Spotify, there are all kinds of playlists that have in some sense a connection to being considered mellow. This corpus will try to take a deeper dive into this, questioning what kind of attributes make a playlist 'mellow', an what don't. This will be done with the use of three different datasets with musical information. These will consists of a 'mellow' playlist, a 'hardcore' playlist, and the playlist for the top 100 songs in the world at the moment. Next, the features and measurement of the playlists will be studied using histograms for loudness and valence, Chromagrams, Tempograms, and a form of clustering. After the analysis, there will be a conclusion section.

### Used data
As said before, the analysis will contain plot for loudness and valence. On top of that, there will be chromagrams and tempograms. In this section, I will briefly delve into the type of data that is used in this portfolio.

The playlists where the data is derived from are 'mellow bars', 'Hardcore mix', and 'top 100 songs global'. For the loudness and valence plots, all tracks will be used in the visualisation. This is the case as well for the dendogram, which can be found at the end of this portfolio. For the tempograms and chromagrams on the other hand, there will be a different approach. For each playlist, 3 songs will be randomly sampled, which will add up to 9 songs in total. Subsequently, there will be 3 rows of 3 chromagrams/tempograms each.

<iframe src="https://open.spotify.com/embed/playlist/37i9dQZF1DWT6MhXz0jw61?utm_source=generator" width="100%" height="380" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>

<iframe src="https://open.spotify.com/embed/playlist/37i9dQZF1EIexrP9IBJXx5?utm_source=generator" width="100%" height="380" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>

<iframe src="https://open.spotify.com/embed/playlist/37i9dQZEVXbNG2KDcFcKOF?utm_source=generator" width="100%" height="380" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>

### Loudness of different styles

```{r, loudness visualisation}
plot1 <- mellow_bars %>%
  plot_ly(x = ~loudness, hoverinfo = "text", text = ~paste("Track:", track.name, "<br>", "Playlist:", playlist_name)) %>%
  add_histogram(name = 'Mellow tracks')

plot2 <- hardcore_tracks %>%
  plot_ly(x = ~loudness, hoverinfo = "text", text = ~paste("Track:", track.name, "<br>", "Playlist:", playlist_name)) %>%
  add_histogram(name = 'Hardcore mix')

plot3 <- top100_tracks %>%
  plot_ly(x = ~loudness, hoverinfo = "text", text = ~paste("Track:", track.name, "<br>", "Playlist:", playlist_name)) %>%
  add_histogram(name = 'Top Songs- Global')

subplot(plot1, plot2, plot3, nrows = 1, shareY = TRUE)%>%
  layout(title = "Loudness of different spotify playlists, captured in an histogram.",
    xaxis = list(title = "Loudness (db)"),
    yaxis = list(title = "Frequency"))
```


***
In this plot, the 'loudness' of three different playlists, 'Mellow bars', 'Hardcore mix', and 'top 100 world' are visualized in histograms. Looking at the chart, one might be surprised to find the loudness of hardcore tracks to be considerably lower in comparison to the Techno and mellow playlists.

### Valence of different styles

```{r, valence visualisation}
plot1 <- mellow_bars %>%
  plot_ly(x = ~valence, hoverinfo = "text", text = ~paste("Track:", track.name, "<br>", "Playlist:", playlist_name)) %>%
  add_histogram(name = 'Mellow tracks')

plot2 <- hardcore_tracks %>%
  plot_ly(x = ~valence, hoverinfo = "text", text = ~paste("Track:", track.name, "<br>", "Playlist:", playlist_name)) %>%
  add_histogram(name = 'Hardcore mix')

plot3 <- top100_tracks %>%
  plot_ly(x = ~valence, hoverinfo = "text", text = ~paste("Track:", track.name, "<br>", "Playlist:", playlist_name)) %>%
  add_histogram(name = 'Top Songs- Global')

subplot(plot1, plot2, plot3, nrows = 1, shareY = TRUE)%>%
  layout(title = "Valence of different spotify playlists, captured in an histogram.",
    xaxis = list(title = "Valence (0-1)"),
    yaxis = list(title = "Frequency"))
```

***
In this plot, the 'Valence' of three different playlists, 'Mellow bars', 'Hardcore mix', and 'top 100 world' are visualized in histograms. A notable insight can be gathered when comparing the plot for mellow tracks and the plot for hardcore tracks. In the plot for mellow tracks, there seems much more tracks with a relatively high valance, in comparison to the hardcore tracks, which have relatively low valence.

### Chromagrams

```{r}
plot <- readRDS("data/chromagrams.RDS")
print(plot)
```

***
Chromagrams of 3 different playlists, each consisting of three songs. More specifically, each row consists of three tracks belonging to a certain playlist. The tracks for the hardcore playlist can be found in the upper row, the tracks for the mellow playlist in the middle row, and the tracks for the top-100 playlist in the last row.

Note: unfortunately, 2 tracks somehow ended up in the wrong place. The song at the end of the row for hardcore songs ("Free nationales - Gangsta"), should be swished with the song at the beginning of the row for mellow songs ("Madball - Bron strong")

Looking at the plots, one could argue that their aren't much differences between the chromagrams of the genre types. In conclution, the kind of keys which are played in a song, does not have a significant inpact on whetehr a track is mellow or not.

### Tempogram
```{r}
tempogram_plots <- readRDS(file = "data/tempogram_plots.RDS")
grid.arrange(grobs = tempogram_plots, ncol = 3)
```

***
Tempograms of 3 different playlists, each consiting of three songs. More specifically, each row consists of three tracks belonging to a certain playlist. The tracks for the top 100 playlist can be found in the upper row, the tracks for the mellow playlist in the middle row, and the tracks for the hardcore playlist in the last row.

Looking at the plot, one can observe that the hardcore songs have almost noi variation in tempo. On the contrary, the mellow songs have a lot of variatoion. Lastly, the songs from the top-100 playlist have a variation that is somewhere in between that of hardcore songs and mellow songs.

In conclusion, one could say that mellow songs have more variation in tempo, in comparison to other genres.

### Dendogram

```{r}
combined_tracks <- rbind(hardcore_tracks, top100_tracks, mellow_bars)
```

```{r}
halloween <-
  mellow_bars |>
  add_audio_analysis() |>
  mutate(
    segments = map2(segments, key, compmus_c_transpose),
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "mean", norm = "manhattan"
      ),
    timbre =
      map(
        segments,
        compmus_summarise, timbre,
        method = "mean"
      )
  ) |>
  mutate(pitches = map(pitches, compmus_normalise, "clr")) |>
  mutate_at(vars(pitches, timbre), map, bind_rows) |>
  unnest(cols = c(pitches, timbre))
```

```{r}
halloween_juice <-
  recipe(
    track.name ~
      danceability +
      energy +
      loudness +
      speechiness +
      acousticness +
      instrumentalness +
      liveness +
      valence +
      tempo +
      duration +
      C + `C#|Db` + D + `D#|Eb` +
      E + `F` + `F#|Gb` + G +
      `G#|Ab` + A + `A#|Bb` + B +
      c01 + c02 + c03 + c04 + c05 + c06 +
      c07 + c08 + c09 + c10 + c11 + c12,
    data = halloween
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(halloween |> mutate(track.name = str_trunc(track.name, 20))) |>
  juice() |>
  column_to_rownames("track.name")
```

```{r}
halloween_dist <- dist(halloween_juice, method = "euclidean")

halloween_dist |> 
  hclust(method = "single") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()
```

***
Dendogram exploring the relations between different tracks of the three used playlists combined.

### Clustering
```{r}
heatmaply(
  halloween_juice,
  hclustfun = hclust,
  hclust_method = "average",  # Change for single, average, or complete linkage.
  dist_method = "euclidean"
)
```

***

### Conclusion and discussion

To answer the question on what makes a track mellow, this corpus have made a couple of visualisations. Of these, the chromagrams, tempograms, valence plots, and loudness plots have been evaluated to help answer the question at hand. In conclution, one can say that mellow tracks have, in comparison to hardcore tracks, more valance, more loudness, and more variation in tempo.

The research have come with a few shortcomings. One of these is a lack of data that have been used in the visualisations. For instance, there were only 10 songs used for each category in the valance and loudness plots. Additionally, the Dendogram and clustering visualisations have not been used in the construction of the conclusion above.
