---
title: "Steven's corpus"
author: "Steven van Wely"
date: "February the 23nd, 2024"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    self_contained: false
    source: https://github.com/Stevgit1/Computational-musicology-homework-7/blob/main/index.Rmd
    theme:
      heading_font:
        google: 
          family: Rajdhani
          wght: 700
      base_font:
        google: Fira Sans
      code_font:
        google: Fira Mono
      bg: "#FFFFFF"
      fg: "#212529" 
      primary: "#2b2bee"
      secondary: "#39d7b8"
      success: "#39d7b8"
      danger: "#fa5577"
      warning: "#ffb14c"
      info: "#0cc7f1"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

```{r, setup}
library(tidyverse)
library(ggplot2)
library(plotly)
library(spotifyr)
library(compmus)
library(ggdendro)
library(heatmaply)
library(tidymodels)
library(cowplot)
```

```{r, setup2}
# adding playlists
mellow_bars <- get_playlist_audio_features("", "37i9dQZF1DWT6MhXz0jw61")
hardcore_tracks <- get_playlist_audio_features("", "37i9dQZF1EIexrP9IBJXx5")
Techno_tracks <- get_playlist_audio_features("", "37i9dQZF1DX6J5NfMJS675")
```

### Mellow: a study

On Spotify, there are all kinds of playlists that have in some sence a connection to being considered mellow. This corpus will try to take a deeper dive into this, questioning what kind of attributes make a playlist 'mellow', an what don't. Specifiaclly, in this version, there will be taken a deeper look at to what extent the attribute 'loudness', as defined by Spotify, has an influence on different song types.
  
### Visualization 1

```{r, first visualization}
plot1 <- mellow_bars %>%
  plot_ly(x = ~loudness, hoverinfo = "text", text = ~paste("Track:", track.name, "<br>", "Playlist:", playlist_name)) %>%
  add_histogram(name = 'Mellow tracks')

plot2 <- hardcore_tracks %>%
  plot_ly(x = ~loudness, hoverinfo = "text", text = ~paste("Track:", track.name, "<br>", "Playlist:", playlist_name)) %>%
  add_histogram(name = 'Hardcore mix')

plot3 <- Techno_tracks %>%
  plot_ly(x = ~loudness, hoverinfo = "text", text = ~paste("Track:", track.name, "<br>", "Playlist:", playlist_name)) %>%
  add_histogram(name = 'Techno bunker')

subplot(plot1, plot2, plot3, nrows = 1, shareY = TRUE)%>%
  layout(title = "Loudness of different spotify playlists, captured in an histogram.")
```


***
In this plot, the 'loudness' of three different playlists, 'Mellow bars', 'Hardcore mix', and 'Techno bunker' are visualized in histograms. Looking at the chart, one might be surprised to find the loudness of hardcore tracks to be considerably lower in comparison to the Techno and mellow playlists. Additionally, the loudness of the Techno tracks and the mellow tracks seem on the same level, which is interesting.

### Conclusion and discussion
In conclusion, hardcore tracks resemble much less 'loudness' then the track types Techno and Mellow. One could however question Spotify's ability to label tracks as loud.

### Tallis plot
```{r tallis}
tallis <-
  get_tidy_audio_analysis("2J3Mmybwue0jyQ0UVMYurH") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)
chapelle <-
  get_tidy_audio_analysis("4ccw2IcnFt1Jv9LqQCOYDi") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)
maria_dist <-
  compmus_long_distance(
    tallis |> mutate(pitches = map(pitches, compmus_normalise, "manhattan")),
    chapelle |> mutate(pitches = map(pitches, compmus_normalise, "manhattan")),
    feature = pitches,
    method = "aitchison"
  )
```

```{r tallis-plot}
maria <-
  maria_dist |>
  mutate(
    tallis = xstart + xduration / 2,
    chapelle = ystart + yduration / 2
  ) |>
  ggplot(
    aes(
      x = tallis,
      y = chapelle,
      fill = d
    )
  ) +
  geom_tile(aes(width = xduration, height = yduration)) +
  coord_fixed() +
  scale_x_continuous(
    breaks = c(0, 60, 105, 150, 185, 220, 280, 327),
    labels =
      c(
        "Ave Maria",
        "Ave cujus conceptio",
        "Ave cujus nativitas",
        "Ave pia humilitas",
        "Ave vera virginitas",
        "Ave preclara omnibus",
        "O Mater Dei",
        ""
      ),
  ) +
  scale_y_continuous(
    breaks = c(0, 45, 80, 120, 145, 185, 240, 287),
    labels =
      c(
        "Ave Maria",
        "Ave cujus conceptio",
        "Ave cujus nativitas",
        "Ave pia humilitas",
        "Ave vera virginitas",
        "Ave preclara omnibus",
        "O Mater Dei",
        ""
      ),
  ) +
  scale_fill_viridis_c(option = "E", guide = "none") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(x = "The Tallis Scholars", y = "La Chapelle Royale")
maria
```

***
New stuff

### new chromogram

```{r pop-chroma}

ewf <-
  get_tidy_audio_analysis("3cfnGXJ9bmiWvFqEO6ff8B?si=fc8b75cc5a254be0") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

loveontop <-
  get_tidy_audio_analysis("1z6WtY7X4HQJvzxC4UgkSf?si=498f16c4cea74323") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

allineed <-
  get_tidy_audio_analysis("164VgxTozx99XCinCB9ITR?si=797ce93143424c36") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)


```

```{r pop-chroma-plots}

ewf_plot <- 
  ewf |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |> 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  geom_vline(xintercept = 53, colour = "red") +
  labs(
    x = "Time (s)", 
    y = NULL, 
    fill = "Magnitude", 
    title = "After the Love Has Gone (Earth, Wind, and Fire)"
  ) +
  theme_minimal() +
  scale_fill_viridis_c() 


loveontop_plot <- 
  loveontop |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |> 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  geom_vline(xintercept = 184, colour = "red") +
  labs(
    x = "Time (s)", 
    y = NULL, 
    fill = "Magnitude", 
    title = "Love On Top (Beyonce)"
  ) +
  theme_minimal() +
  scale_fill_viridis_c() 


allineed_plot <- 
  allineed |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |> 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  geom_vline(xintercept = 137, colour = "red") +
  geom_vline(xintercept = 186, colour = "red") +
  labs(
    x = "Time (s)", 
    y = NULL, 
    fill = "Magnitude", 
    title = "All I Need (Collier, Mahalia, Ty Dolla $ign)"
  ) +
  theme_minimal() +
  scale_fill_viridis_c() 

plot_grid(ewf_plot, loveontop_plot, allineed_plot, ncol = 1)
```
***
Nice stuff

### New plot

```{r twenty-five}
circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)
key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

get_tidy_audio_analysis("4EwsiiT0hwIi38gAVEEtIV") |>
  compmus_align(sections, segments) |>
  select(sections) |>
  unnest(sections) |>
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "acentre", norm = "manhattan"
      )
  ) |>
  compmus_match_pitch_template(key_templates, "aitchison", "manhattan") |>
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(option = "E", guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "", fill = "Distance")
```

***
New visualisation

### Tempogram
```{r}
europapa_url <- '0uHrMbMv3c78398pIANDqR'
graveola_url <- '6PJasPKAzNLSOzxeAH33j2'
graveola <- get_tidy_audio_analysis(europapa_url)

graveola |>
  tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```

***
New tempogram

### Dendogram
```{r}
joost_url = '3HoBmXOKHexP12ekdvXjLY'
halloween <-
  get_playlist_audio_features("Joost Klein", joost_url) |>
  add_audio_analysis() |>
  mutate(
    segments = map2(segments, key, compmus_c_transpose),
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "mean", norm = "manhattan"
      ),
    timbre =
      map(
        segments,
        compmus_summarise, timbre,
        method = "mean"
      )
  ) |>
  mutate(pitches = map(pitches, compmus_normalise, "clr")) |>
  mutate_at(vars(pitches, timbre), map, bind_rows) |>
  unnest(cols = c(pitches, timbre))
```

```{r}
halloween_juice <-
  recipe(
    track.name ~
      danceability +
      energy +
      loudness +
      speechiness +
      acousticness +
      instrumentalness +
      liveness +
      valence +
      tempo +
      duration +
      C + `C#|Db` + D + `D#|Eb` +
      E + `F` + `F#|Gb` + G +
      `G#|Ab` + A + `A#|Bb` + B +
      c01 + c02 + c03 + c04 + c05 + c06 +
      c07 + c08 + c09 + c10 + c11 + c12,
    data = halloween
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(halloween |> mutate(track.name = str_trunc(track.name, 20))) |>
  juice() |>
  column_to_rownames("track.name")
```

```{r}
halloween_dist <- dist(halloween_juice, method = "euclidean")

halloween_dist |> 
  hclust(method = "single") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()
```

***
New Dendogram

### Clustering
```{r}
heatmaply(
  halloween_juice,
  hclustfun = hclust,
  hclust_method = "average",  # Change for single, average, or complete linkage.
  dist_method = "euclidean"
)
```

***
New stuff